Cluster Difference Imaging Photometric Survey (CDIPS) II. TOI$\,$837: A Validated Jupiter-Sized Planet in IC~2602

### 2020-07-02 paper/ms.tex
update manuscript

%
% TODO:
% * Zhou or Quinn: what is the SpType? Add to abstract, tables, and
% text. G0V has been assumed.
%

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf TESS light curve of \tn\ from Sectors 10 and 11.} {\it Top}:
    \texttt{PDCSAP} median-subtracted relative flux at 2-minute
    sampling in units of parts-per-thousand ($\times 10^{-3}$).
    Spot-induced stellar variability is the dominant signal; flares are
    shown with gray crosses.  Dashed
    lines indicate the five transits observed by TESS.  {\it Bottom}:
    Zoomed windows of individual transits.
    Gray lines are the best-fit transit model, which includes a local
    quadratic trend in each transit window.
		\label{fig:rawzoom}
	}
\end{figure*}

@@

The TESS project detected the transits in the 2-minute data and the
community was alerted on 17 June 2019.  While our CDIPS FFI light
curves also show the transits, the 2-minute data have better sampling
cadence, and so we opted to use the Presearch Data Conditioning (PDC)
light curve with the default aperture for our
analysis~\citep{smith_kepler_2012,stumpe_multiscale_2014,jenkins_tess_2016,smith_finding_2016}.

Figure~\ref{fig:rawzoom} shows the data.  The dominant modulation
induced by starspots coming into and out of view has a peak-to-peak
amplitude of about 2.3\%, and a period of about 3 days.  The dips
suggestive of a grazing transiting planet recur roughly every 8 days,
and have a depth of about 0.4\%.  A few flares are visible.  A
phase-folded view of the TESS transits with starspot variability
removed is shown in Figure~\ref{fig:jointphot}, and is described in
depth in Section~\ref{subsec:planet}.

@@

%FIXME: this paragraph doesn't really belong here?
We examined the CDIPS FFI light curves of the target, which are
available on MAST \citep{bouma_cluster_2019}. Three light curves are
available, based on photometric apertures with a radius of 1, 1.5, or
2.5 pixels.  In the raw difference-image light-curves, as well as the
PCA-detrended light-curves, dips of depth $\approx$0.35\% are visible,
and their properties do not significantly vary with aperture size.
Within $\approx20$'', the dips are consistent with originating from
the target star.

@@

{\bf Carl TODO: please describe and correct as appropriate}
The observations of \tn\ were initially acquired by
\citet{ziegler_soar_2020} as part of the Southern Astrophysical
Research (SOAR) TESS Survey using the High Resolution Camera (HRCam;
\citealt{tokovinin_ten_2018}).  The HRCam $I$-band filter is described
by \citet{tokovinin_ten_2018}.  The points in Figure~\ref{fig:followup}
show the resulting measured 5-$\sigma$ detectable contrasts.  The
lines are linear smoothing fits between the regimes of the diffraction
limit, the ``knee'' at $\approx 0.2''$, and the slow decrease until
$\approx 1.5''$, beyond which the speckle patterns become
de-correlated.  Star A (TIC 847769574) was detected at the expected
location and brightness contrast, and no additional companions were
found.

@@

{\bf Phil + Karen TODO: please correct as appropriate}
We scheduled the observations using {\bf \texttt{SOFTWARE}} (CITE).
The photometric data were calibrated and extracted using the
\texttt{AstroImageJ} software package
\citep{collins_astroimagej_2017}.  Comparison stars of similar
brightness were used to produce the final light curves, each of which
showed a roughly 4$\,$ppt dip near the expected transit time.

\paragraph{Custom aperture analysis}
Based solely on the TESS data, both Star A and Star B
were possible sources of blended eclipsing binary signals.
For instance, a 30\% eclipse of Star A could produce
dips with the correct depth ($\approx$0.4\%) to explain the \tn\ signal.
Star B is resolved in the ground-based images; Star A is not.

@@

To assess the possibility that Star A is the eclipsing body,
we created light curves with a custom set of circular apertures
with radii of $2.1''$ and positions ranging from Star A (2.3$''$ west of \tn) to 2.3$''$
east of \tn.  We did not detect any variation of the transit depth
along this line of light curves, as would be expected if Star A were
the eclipse host.  The apertures east of \tn\ exclude over 90\% of the flux from Star A.
The eclipse on Star A would therefore need to be excessively deep to
produce the observed eclipse depth.
We take this as evidence that \tn\ is the source of the transit signal
to within $\approx2.0''$.

@@

{\bf Amaury TODO: please correct as appropriate}
We observed {\bf N} transits with the 0.40m ASTEP 400 telescope at
Concordia Station, located at Dome C on the Antarctic Plateau
\citep{daban_astep_2010}.  ASTEP 400 has a dichroic that sends blue
wavelengths to a guiding camera, and red wavelengths to a filter-free
science camera with wavelength sensitivity roughly akin to R band.
The observations were acquired on the nights of {\bf DATE 1} and {\bf
DATE 2}.  The outdoor temperature was a balmy {\bf XX$^\circ$}.  The
data were reduced using {\bf XX} software, and showed the transit
roughly at the expected time and depth.

@@

{\bf George TODO: please describe the formal secondary line search in
more depth in this paragraph} We searched for secondary lines in the
CHIRON spectra by cross-correlating against a {\bf template of some
kind}.  This rules out lines with flux ratios of $\approx 2\%$, if the
lines are well-separated in velocity.  To derive the star's physical
parameters, we cross-correlated the template against {\bf something},
and used the {\bf \texttt{nameofcode}}.  We measured the radial
velocities through a least-squares deconvolution of the observed
spectra against the same synthetic template.  The resulting velocities
are given in Table~\ref{tab:rvs}, and shown in Figure~\ref{fig:followup}.

@@

We used the dedicated instrument pipeline to extract 1-dimensional,
wavelength calibrated spectra.  To measure radial velocities, {\bf
Rafael TODO: what procedure was used?  What software packages?}.  The
velocities are given in Table~\ref{tab:rvs}, and shown in
Figure~\ref{fig:followup}.

@@

% results from 20200624_simple_fpp_weak_sec 
We ran \texttt{VESPA}\footnote{We used \texttt{VESPA-0.6} and
\texttt{isochrones-1.2.2}.}, and directly incorporated our constraints
of the SOAR $I$-band contrast curve and a non-detection of secondary
eclipses with a depth set at roughly twice the limits from the SPOC
vetting report (0.1\%).  We verified that changing the secondary
eclipse depth limit did not significantly affect the results.  We set
the maximum aperture radius at $2''$, based on our ground-based
photometry.  Incorporating the constraints from
Figure~\ref{fig:fpscenario}, our nominal false positive probability
analysis excluded EB and HEB scenarios.  This yielded a FPP of 0.21\%
for \pn, sufficient for formal validation as a planet
\citep{morton_efficient_2012}.  We did not incorporate our constraint
that \tn\ is not double-lined, which rules out an additional portion
of BEB parameter space.  Had we not acquired multicolor ground-based
photometry, and been unable to exclude HEB scenarios, the FPP would
have risen to 8\%.  However, since the transits are achromatic
(Figure~\ref{fig:jointphot}), particularly in Johnson-B band, we can
rule out HEB scenarios.

@@

We derived physical parameters of \pn\ based on the photometry, we
fitted the TESS and ground-based transits simultaneously.
We trimmed the data to

\subsubsection{Transit Fitting}
Simultaneous dip plus rotation period (GP) fit. Use celerite plus PyMC3/exoplanet.
Mention FPP?

%TODO: describe 0.3-day robust Huber spline (or whatever detrending is
%eventually used)

\subsubsection{Comments on Joint Transit and Radial Velocity Fitting}

The available data for \tn\ put us in an odd position. The transit
data provide a minimum size due to the grazing geometry; the radial
velocity data provide a maximum mass ($\approx$1.20 $M_{\rm jup}$).
We could combine the two with an empirically-calibrated planetary
mass-radius relation to force a more restrictive upper bound on the
planetary size \citep[{\it e.g.},][]{chen_probabilistic_2016}.  The
result of such an exercise would tautologically be that the planet
would be smaller than the largest known Jovian planets ($\approx
2R_{\rm Jup}$).  This would also imply that the impact parameter is
slightly below unity.  The planet's size, in our eyes, would remain
poorly constrained even if we were to do this exercise. {\bf So we
don't do it.}

\subsubsection{multicolor transit fits}
Single transit at a time, with limited

\subsubsection{Additional Transits}
None of the extra dips in the PDCSAP light-curve (e.g., ``up-down''
spikes at BTJD 1572 and 1601) seem likely to be planetary
We checked that i) they were not present in the SAPFLUX light-curves,
and ii) that they were not present in the CDIPS light-curves (either
raw, or PCA-detrended).

To make this quantitative, we did injection-recovery.

@@

% \input{WASP-4b_transit_time_table.tex}
% \input{WASP-4b_rv_table.tex}
% \input{model_fit_table.tex}
% \input{rv_model_posterior_table.tex}
% \input{pdot_table.tex}

\clearpage
\input{TOI837_phot_table.tex}
\input{TOI837_rv_table.tex}
%\input{model_comparison_table.tex}
\clearpage

>>>>

%
% TODO:
% * Zhou or Quinn: what is the SpType? Add to abstract, tables, and
% text. G0V has been assumed.
% * remove all bf todo's
%

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.9\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf TESS light curve of \tn\ from Sectors 10 and 11.} {\it Top}:
    \texttt{PDCSAP} median-subtracted relative flux at 2-minute
    sampling in units of parts-per-thousand ($\times 10^{-3}$).
    Spot-induced stellar variability is the dominant signal; flares are
    shown with gray crosses.  Dashed
    lines indicate the five transits observed by TESS.  {\it Bottom}:
    Zoomed windows of individual transits.
    Gray lines are the best-fit transit model, which includes a local
    quadratic trend in each window.
		\label{fig:rawzoom}
	}
\end{figure*}

@@

The TESS project detected the transits in the 2-minute data and the
community was alerted on 17 June 2019.  While our CDIPS FFI light
curves also showed the transits, the 2-minute data had better sampling
cadence, and so we opted to use the Presearch Data Conditioning (PDC)
light curve with the default aperture for our
analysis~\citep{smith_kepler_2012,stumpe_multiscale_2014,jenkins_tess_2016,smith_finding_2016}.

Figure~\ref{fig:rawzoom} shows the data.  The dominant modulation
induced by starspots coming into and out of view has a peak-to-peak
amplitude of about 2.3\%, and a period of about 3 days.  The dips
suggestive of a grazing transiting planet recur roughly every 8 days,
and have a depth of about 0.4\%.  A few flares are visible.  A
phase-folded view of the TESS transits with starspot variability
removed is shown in Figure~\ref{fig:jointphot}. Our fitting procedure
for the models shown in these plots is described in
Section~\ref{subsec:planet}.

@@

% %FIXME: this paragraph doesn't really belong here?
% We examined the CDIPS FFI light curves of the target, which are
% available on MAST \citep{bouma_cluster_2019}. Three light curves are
% available, based on photometric apertures with a radius of 1, 1.5, or
% 2.5 pixels.  In the raw difference-image light-curves, as well as the
% PCA-detrended light-curves, dips of depth $\approx$0.35\% are visible,
% and their properties do not significantly vary with aperture size.
% Within $\approx20$'', the dips are consistent with originating from
% the target star.

@@

\edit1{Carl TODO: please describe and correct as appropriate}
The observations of \tn\ were initially acquired by
\citet{ziegler_soar_2020} as part of the Southern Astrophysical
Research (SOAR) TESS Survey using the High Resolution Camera (HRCam;
\citealt{tokovinin_ten_2018}).  The HRCam $I$-band filter is described
by \citet{tokovinin_ten_2018}.  The points in Figure~\ref{fig:followup}
show the resulting measured 5-$\sigma$ detectable contrasts.  The
lines are linear smoothing fits between the regimes of the diffraction
limit, the ``knee'' at $\approx 0.2''$, and the slow decrease until
$\approx 1.5''$, beyond which the speckle patterns become
de-correlated.  Star A (TIC 847769574) was detected at the expected
location and brightness contrast, and no additional companions were
found.

@@

\edit1{Phil + Karen TODO: please correct as appropriate}
We scheduled the observations using {\bf \texttt{SOFTWARE}} (CITE).
The photometric data were calibrated and extracted using the
\texttt{AstroImageJ} software package
\citep{collins_astroimagej_2017}.  Comparison stars of similar
brightness were used to produce the final light curves, each of which
showed a roughly 4$\,$ppt dip near the expected transit time.

\paragraph{Custom aperture analysis}
Based solely on the TESS data, both Star A and Star B
were possible sources of blended eclipsing binary signals.
Star B is resolved in the ground-based images; Star A is not.

@@

To assess the possibility that Star A is an eclipsing body,
we created light curves with a custom set of circular apertures
with radii of $2.1''$ and positions ranging from Star A (2.3$''$ west of \tn) to 2.3$''$
east of \tn.  We did not detect any variation of the transit depth
along this line of light curves, as would be expected if Star A were
the eclipse host.  The apertures east of \tn\ exclude over 90\% of the flux from Star A.
The eclipse on Star A would therefore need to be excessively deep to
produce the observed eclipse depth.
We take this as evidence that \tn\ is the source of the transit signal
to within $\approx2.0''$.

@@

\edit1{Amaury TODO: please correct as appropriate}
We observed {\bf N} transits with the 0.40m ASTEP 400 telescope at
Concordia Station, located at Dome C on the Antarctic Plateau
\citep{daban_astep_2010}.  ASTEP 400 has a dichroic that sends blue
wavelengths to a guiding camera, and red wavelengths to a filter-free
science camera with wavelength sensitivity roughly akin to R band.
The observations were acquired on the nights of {\bf DATE 1} and {\bf
DATE 2}.  The outdoor temperature was a balmy {\bf XX$^\circ$}.  The
data were reduced using {\bf XX} software, and showed the transit
roughly at the expected time and depth.

@@

\edit1{George TODO: please describe the formal secondary line search in
more depth in this paragraph} We searched for secondary lines in the
CHIRON spectra by cross-correlating against a {\bf template of some
kind}.  This rules out lines with flux ratios of $\approx 2\%$, if the
lines are well-separated in velocity.  To derive the star's physical
parameters, we cross-correlated the template against {\bf something},
and used the {\bf \texttt{nameofcode}}.  We measured the radial
velocities through a least-squares deconvolution of the observed
spectra against the same synthetic template.  The resulting velocities
are given in Table~\ref{tab:rvs}, and shown in Figure~\ref{fig:followup}.

@@

We used the dedicated instrument pipeline to extract 1-dimensional,
wavelength calibrated spectra.  To measure radial velocities, \edit1{
Rafael TODO: what procedure was used?  What software packages?}.  The
velocities are given in Table~\ref{tab:rvs}, and shown in
Figure~\ref{fig:followup}.

@@

% results from 20200624_simple_fpp_weak_sec 
We ran \texttt{VESPA}\footnote{We used \texttt{VESPA-0.6} and
\texttt{isochrones-1.2.2}.}, and directly incorporated our constraints
of the SOAR $I$-band contrast curve and a non-detection of secondary
eclipses with a depth set at roughly twice the limits from the SPOC
vetting report (0.1\%).  We verified that changing the secondary
eclipse depth limit did not significantly affect the results.  We set
the maximum aperture radius at $2''$, based on our ground-based
photometry.  Incorporating the constraints from
Figure~\ref{fig:fpscenario}, our nominal false positive probability
analysis excluded EB and HEB scenarios.  This yielded an FPP of 0.21\%
for \pn, sufficient for formal validation as a planet
\citep{morton_efficient_2012}.  We did not incorporate our constraint
that \tn\ is not double-lined, which rules out an additional portion
of BEB parameter space.  Had we not acquired multicolor ground-based
photometry, and been unable to exclude HEB scenarios, the FPP would
have risen to 8\%.  However, since the transits are achromatic
(Figure~\ref{fig:jointphot}), particularly in Johnson-B band, we can
rule out HEB scenarios.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f10.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{ 
		{\bf Posterior probabilities of
			impact parameter, planet-to-star size ratio, and stellar density.} The photometric
		data fitted in Section~\ref{subsec:planet} allow for
		nonphysical planet sizes ($R_{\rm p}>2R_{\rm Jup}$;
		$\log(R_{\rm p}/R_\star) \gtrsim -1.6$) due to a classic degeneracy
		between the planet size and impact parameter.  Non-planetary sizes
		would however imply non-planetary masses, which are ruled out by our
		radial velocity measurements.
		This plot was made using \texttt{corner} \citep{corner_2016}.
		\label{fig:subsetcorner}
	}
\end{figure}

We fitted the available time-series photometry of \pn\ as follows.
For the PDCSAP light curve, we first eliminated points that had
quality flags corresponding to any of bits $\{3, 4, 6, 8, 11, 12\}$.
This excluded cadences affected by coarse spacecraft pointing,
reaction wheel desaturation events, manual flags, cosmic ray hits, and
straylight from the Earth or Moon being present.  Inspecting the data,
we also manually excluded the two flares shown in
Figure~\ref{fig:rawzoom}.  We then trimmed the TESS data to windows of
$\pm 7\,{\rm hr}$ centered on each transit.

Our joint model for the TESS and ground-based data was a
\citet{exoplanet:agol19} transit with physical and orbital parameters
shared across all transit windows, plus a local quadratic trend
allowed within each window.  The parameters and priors
are listed in Table~\ref{tab:posterior}.  In brief, we fitted for the
shared stellar parameters ($\log g, R_\star, u_0, u_1$), and the
shared planetary parameters ($t_0, P, b, \log(R_{\rm p}/R_\star)$).
There were also three free trend parameters for each of the nine
transits to account for the rotational variability.  This yielded a
total of 35 free parameters, of which 8 have physical relevance, and
27 are nuisance parameters.

We fitted the model using \texttt{PyMC3}
\citep{salvatier_2016_PyMC3,exoplanet:theano}. 
For the exoplanet transit, we used the \texttt{exoplanet} code
\citep{exoplanet:exoplanet}.  
After initializing each model with the parameters of the
maximum {\it a posteriori} model, we assumed a Gaussian
likelihood, and sampled using \texttt{PyMC3}'s gradient-based
No-U-Turn Sampler \citep{hoffman_no-u-turn_2014}. We used $\hat{R}$
as our convergence diagnostic \citep{gelman_inference_1992}.

While alternative approaches are available for simultaneously fitting
space and ground based photometry (in addition to radial velocimetry),
we opted for this approach because {\it i)} the radial velocitmetry on
its own did not show evidence for a planetary signal, and {\it ii)} we
expected that this approach would provide the best measurement of the
planetary ephemeris, which is key for follow-up efforts.  The
assumption of a constant radius across all bandpasses was tested by
independently fitting each ground-based transit while letting the
planetary radius float (Section~\ref{subsec:colorphot}). The transit
depths did not significantly change between different bandpasses.

The posteriors from this procedure are given in
Table~\ref{tab:posterior}.  Crucially important is the point that the
transit is grazing.  The grazing geometry hinders our ability to infer
the stellar density and planet-to-star size ratio. The relevant
posterior probabilities are shown in Figure~\ref{fig:subsetcorner}.

The available data for \tn\ put us in an odd position. The transit
data provide a minimum size due to the grazing geometry; the radial
velocity data provide a maximum mass ($\approx$1.20 $M_{\rm jup}$).
We could combine the two with an empirically-calibrated planetary
mass-radius relation to force a more restrictive upper bound on the
planetary size \citep[{\it e.g.},][]{chen_probabilistic_2016}.  The
result of such an exercise would tautologically be that the planet
would be smaller than the largest known Jovian planets ($\approx
2R_{\rm Jup}$).  This would also imply that the impact parameter is
slightly below unity.  The planet's size, in our eyes, would remain
poorly constrained even if we were to do this exercise, so we opt to 
fit the photometry and radial velocities separately, to ensure clarity
about which datasets are providing which constraints.

% TODO: decide whether to include this. it's not necessary!
% \subsubsection{Additional Transits}
% None of the extra dips in the PDCSAP light-curve (e.g., ``up-down''
% spikes at BTJD 1572 and 1601) seem likely to be planetary
% We checked that i) they were not present in the SAPFLUX light-curves,
% and ii) that they were not present in the CDIPS light-curves (either
% raw, or PCA-detrended).
% To make this quantitative, we did injection-recovery(?)

@@

\clearpage
\input{TOI837_phot_table.tex}
\input{TOI837_rv_table.tex}
\clearpage
